\framework: Morphology-based Entity and Relational Entity Extraction Framework for Arabic

### 2018-04-20 tex/merf/TALMajor/answer/Reviewer1.comments.tex
done till 3.19

\answer{
  We clarified the following notations.... 
  The new notations appear in Section AA in the paper. 
}
\itodo{
Pending...
}

@@

\answer{
We modified the description of isA to address this issue. 
Basically, isA is used to detect the `exact match' between a user-chosen value for a morphological feature and the corresponding feature value in a morphological solution of a word.
}

\done{
Add two sentences in the paper to say what isA does.\\ 
Changes are highlighted in red at beginning of Subsection~\ref{subsec:grammar}.
}

@@

\answer{
Manual annotators inspected the outcome of the tools and provided 
corrections where tools made mistakes. 
The corrections form the manual gold annotation that we
compared against. 
We indicate that in the Results Section. 
}
\done{
  Add the above to the results section.\\ 
Done.
}

@@

\answer{
  \framework does not require training. 
  It is an interactive tool where the user provides the formulae, expressions,
    relation tuples and the associated code. 
  The user inspects the results and refines the model accordingly. 
  The testing set is available with the tool. 
}

\itodo{
  Ameen will get the size from the files and provide. 
}

@@

\answer{
A research assistant 6 hours course work and 14 hours 
teaching duties performed the task in two months. 
We clarify this issue in the results Section.
}

\done{
Add this to the results/evaluation.\\
Done
}

@@

\answer{
\framework is open source and available online. We modified the
contributions in the Introduction Section to indicate that. 
}

>>>>

\answer{
  We added a methodology section that precedes the formal 
    notation. 
  We also embedded a running example along with the formal
    notations. 
}

@@

\answer{
  We clarified the \cci{isA} predicate in Section~\ref{subsec:grammar}.
  Basically, isA is used to detect the `exact match' 
  between a user-chosen value for a morphological feature and the 
  corresponding feature value in a morphological solution of a word.

  We also replaced a mistake that referenced the \cci{isA} predicate
  in Figure~\ref{fig:intromotiv} by the correct predicate \cci{isSyn}. 
}

@@

\answer{
  We added description of the case studies in the Results Section~\ref{sec:results}. 
  The reference chunks are manually created gold annotations.
  Manual annotators inspected the outcome of the tools and provided 
corrections where tools made mistakes. 
The corrections form the manual gold annotation that we
compared against. 
We indicate that in the Results Section. 
}

@@

\answer{
  We added a column to Table~\ref{tab:results} to refer to the size 
  of the annotated data for each case study and we discussed the numbers
  in the Results Section~\ref{sec:results}. 
  \framework does not require training. 
  It is an interactive tool where the user provides the formulae, expressions,
    relation tuples and the associated code. 
  The user inspects the results and refines the model accordingly. 
  The testing set is available with the tool. 
}

@@

\answer{
  We clarify this issue in the Results Section~\ref{sec:results}.
A research assistant 6 hours course work and 14 hours 
teaching duties performed the task in two months. 
}

@@

\answer{
\framework is open source and available online. We modified the
contributions in the Introduction Section~\ref{sec:introduction} to indicate that. 
}

### 2018-04-20 tex/merf/TALMajor/answer/Reviewer2.comments.tex
done till 3.19

\item \label{Review.2.1} 
The conceptual and implementation aspects of the framework 
were introduced in not clear and differentiated form. 
The authors are advised to separate these issues in formal 
definitions and to present the resulted software tool in 
terms of more clear software techniques.
\answer{
We introduced the following formal definitions...

We clarify the implementation in the following manner. 
Tranformative algorithms: 1. RE to FSM
                          2. Entities and relations to garphs
Generative algorithms: 1. Actions to code
... 
}

@@

\answer{
  Section XXX discussed morphological analysis in Arabic. 
  We extend that in ...
  We highlight that by referring to the Section in the 
  Introduction. 
}

\item \label{Review.2.3} 
Finally, I believe this framework should be also expressed 
in terms of the complexity and decidability problems, 
particularly when evaluating the required time of different 
development tasks. 
\answer{
  The problem is decidable: RE to FSM. 
  The problem might be not tractable: NDFSM to FSM might 
  be exponential. 
  In practice Table XXX provides runtime results... 
  We modified the paper to include the above. 
}

\item \label{Review.2.4} 
Providing the system with GUI is an advantage, 
However the system requires probably to some extend 
good linguistic and formal languages expertise to optimize 
the benefit of the implements tool.
\answer{
We discuss the target audience of the MERF tool in Section XXX. 
We prepared a set of 45 minutes slides based on the regular
expression chapter in Martin Jurafsky's book to... 
}

>>>>

\item \label{Review.2.1} 
The conceptual and implementation aspects of the framework 
were introduced in not clear and differentiated form. 
The authors are advised to separate these issues in formal 
definitions and to present the resulted software tool in 
terms of more clear software techniques.

\answer{
We introduced a Methodology Section~\ref{sec:methodology} to separate
the conceptual aspects from the implementation aspects of the
framework.

We refer to the collection of regular expressions as a grammar, 
we also provide information about the data model used to store
the user defined tag types. 
In the results Section~\ref{sec:results} we discuss the 
complexity of the simulators.

We now relate the action code API to corresponding software design patterns in Section~\ref{sec:relations-actions}, 
%
We discuss the translation from grammar to NDFSA. We also 
discuss the generative nature of the action code as a shared 
object library. 
}

@@

\answer{
  We provide a Methodology Section~\ref{sec:methodology} 
    that benefits directly from the preceding Background.
  We introduced new Figure~\ref{f:overview} to show that 
  the morphological analysis is the first preprocessing 
  step in the data flow of \framework. 
  Section~\ref{sec:morph} that discusses morphological 
    analysis with solution examples in Table~\ref{tab:samplesolution}.
  Section XXX discussed morphological analysis in Arabic. 
  We extend that in ...
  We highlight that by referring to the Section in the 
  Introduction. 
}

\item \label{Review.2.3} 
Finally, I believe this framework should be also expressed 
in terms of the complexity and decidability problems, 
particularly when evaluating the required time of different 
development tasks.

\answer{
  We added discussion in the Results Section~\ref{sec:results}
    to discuss and comment on the complexity of the 
    regular expression to non-deterministic finite state
    automata (NDFSA), and the simulation of the NDFSA  
    with the morphology based Boolean formulae tags. 
}

\item \label{Review.2.4} 
Providing the system with GUI is an advantage, 
However the system requires probably to some extend 
good linguistic and formal languages expertise to optimize 
the benefit of the implements tool.

\answer{
We added a statement at the end of the Discussion 
Subsection~\ref{sec:discuss} to reflect the comment of the reviewer. 
}

### 2018-04-20 tex/merf/TALMajor/answer/Reviewer3.comments.tex
done till 3.19

\answer{
  We added a background section with definitions... 
}

\done{
Pending...
}

@@

\answer{
  Thank you. We added a tutorial/video demo... 
}

@@

\answer{
  We added a methodology section...
}

\itodo{
  We should the following sections.
    1. \framework methodology: diagram with its description, and the 
    reference to the example +...
    2. \framework supporting tools: the current \framework section
    3. \framework GUI: remain as is

  Look for high level stuff in the (2) and (3) sections that can move
    into (1).

    When this is done go over the answers to make sure the references
    to answers in the paper are accurate.
}

@@

\answer{
  We added subsection Results.X to discuss the evaluation 
  strategy. 
  We also added extended the discussion section... 
}

\itodo{same answer from Reviewer 1}

@@

\answer{
  Thank you. We combine both as motivation and we include that
  as a subsection of the introduction. 
}

\done{
  We combine the Motivation section with \framework section (Now Section~\ref{sec:framework}).
}

@@

\answer{
  We rephrased the sentence in the paper to read as follows. 
  \begin{quote}
The leaf nodes of the match trees are matches to formulae and the internal nodes represent roots to sub-expression matches.
  \end{quote}
}

@@

\answer{
  We use formula to define predicates that range over 
  the morphological features. 
  We consider the matches of the formulae as tokens/tags (maybe
  we call them tag types.)
  We then use regular expressions (maybe we should use rules).

  We modify the paper by adopting the suggested terminology: 
  rule for regular expression and parse tree for match tree. 
}

\itodo{
Lets use regular grammar and introduce that with similarity to antler.
MRE becomes a macro maybe a good acronym is morphology based regular grammar rule 
MRGR. 
}

@@

\answer{
  We introduced a new Methodology Section~\ref{sec:methodology} and
  we moved the content of old Section 2 to it. 
}

@@

\answer{
We split Figure 3 (the flow diagram figure) 
into three Figures and provide explanations
for each. 
a. Arabic Text (Consider MBF as input coming from another figure) ----> Tags.
b. Rest of the figure... 
c. RTC + Diff. + Visualizatiin 
}

\itodo{
  Edit the figure and split it  and refer to the (a), (b) and (c) in the paper. 
}

\item \label{Review.3.13}
b) Syn$^k$ is nicely explanied but I did not recognize it as 
your own idea. if this is your idea to use English translation 
as a pivot to extract Arabic synonyms then it needs to have 
more focus because I think this is new.

@@

\answer{
F is the set of morphological features where each element is in turn a set of the values of that feature.
Hence, A is one such set corresponding to a morphological feature, and CF is a feature value in A.
We clarify this point in Subsection~\ref{subsec:grammar}.
}

>>>>

\answer{
  We cleared notations in the paper that might have led to 
    the confusion.
    For example, we cleared (1) the notation for morphological solutions,
    (2) the use of symbol $R$ as a both a relation tuple and a 
    tag set. 
    We clarified the \cci{isA} and \cci{contains} predicates. 
    We define the power set notation $2^A$ where $A$ is a set
    and $2^A$ is the set of all subsets of $A$.
}

@@

@@

\answer{
  We added a methodology Section~\ref{sec:methodology} with a 
    new Figure~\ref{f:overview} that illustrates the four
    processes involved in the \framework Methodology. 
}

@@

\answer{
  We added more discussions to the results Section~\ref{sec:results}
    including case study size, development time evaluation, 
    complexity, and a discussion on required expertise. 
}

@@

\answer{
  Thank you for the note. 
  We removed the motivation Section and embedded its content as 
  a running example in the \framework Components Section~\ref{sec:tool}. 
}

@@

\answer{
  We rephrased the sentence in the paper to read as follows. 
  \begin{quote}
The leaf nodes of the trees are matches to formulae and the internal nodes represent roots to sub-expression matches.
  \end{quote}
}

@@

\answer{
  We modified the paper to refer to the collection of the regular
  expressions as a local grammar. 
  We also use matching parse trees instead of match trees in the paper now. 
  We added a sentence in the Introduction Section to 
  clarify that ``formula'' across the paper denotes a morphology
  based Boolean formula (MBF). 
  We do the same for regular expressions and rules. 
}

@@

\answer{
  We removed the Motivation Section (2 in the old version) and
  embedded its contents in the \framework Components Section (4 in the
  old version and \ref{sec:tool} now). 
  We also introduced a Methodology Section~\ref{sec:methodology}. 
}

@@

\answer{
  We replaced the figure with Figure~\ref{f:overview}.
  Figure~\ref{f:overview} is 
  a clearer and is composed of four
  separate processes. 
  We discuss the new Figure in the Methodology Section~\ref{sec:methodology}.
}

\item \label{Review.3.13}
b) Syn$^k$ is nicely explained but I did not recognize it as 
your own idea. if this is your idea to use English translation 
as a pivot to extract Arabic synonyms then it needs to have 
more focus because I think this is new.

@@

\answer{
F is a sets of sets. It is a set of morphological feature sets.
Each morphological feature set is in turn a set of values of that feature.
CF belongs to A denotes that CF is a value of a morphological feature
set  A. 
We clarify this point in SubSection~\ref{subsec:grammar}.
}

### 2018-04-20 tex/merf/TALMajor/gui.tex
done till 3.19

\framework saves the tags and the tag types in a user friendly format using
the JavaScript Object Notation (JSON) format~\cite{nolan2014javascript}.
The tag file keeps the paths of the separate text and tag type files. 
It also contains a list of tags with their tag type identifiers,
position in text, length, and word index.
The tag type file contains MBF and MRE tag types. 
%Each object has an array value containing the user-defined tag types. 
Each tag type contains the name, description, expression, and the visualization legends.

@@

\begin{figure}[tb!]
  \centering
  \includegraphics[width=\textwidth]{figures/treegraph}
  \caption{MRE annotated Text, MRE match tree, and entity-relation graph}
  \label{fig:treegraph}
\end{figure}

@@

The match tree view shows the selected match in a graph view.
Figure~\ref{fig:treegraph} shows the MRE match tree of the direction task 
\RL{dby mwl `l_A mqrbT mn h_dA Almbn_A}(Dubai Mall is located near this building). 
%Leaf nodes show the match text, internal nodes show the operations, and edge labels shows the MBF tag type name.

@@

A snapshot of the GUI in 
Figure~\ref{fig:treegraph} shows in an interactive graph view
the entity-relation graph of the match of the user defined relation 
extracted from the match tree of the MRE.

In the computational action editor, an advanced user can 
enter C++ code and use the \framework API to program and process 
subexpression matches. 
%RE matches In order to add computational actions to an MRE sub-expression, 
%the user selects to edit the actions in the MRE tag type editor. 
%The view allows the user to specify the actions in a user friendly manner. 
%\framework provides the user with an API to access the sub-expression match features easily and define the pre-match actions. 
%The match features include the text, position, length, and numerical value if applicable, and morphological features for MBF matches. 
%\framework also enables the user to add C++ declarations and library includes.

@@

In the analysis view, the user provides 
two tag sets $R_1$ and $R_2$ and 
two tag type sets ${\cal T}_1$ and ${\cal T}_2$ as input. 
%The analysis view provides views of the differences between the tag types and 
%the tag sets.
%
The tag type difference view shows the text annotated in three panes: 
1) the common tag types ${\cal T}_1 \cap {\cal T}_2$,
2) the tag types in ${\cal T}_1$ but not in ${\cal T}_2$, 
and 3) the tag types in ${\cal T}_2$ and not in ${\cal T}_1$.

Similarly, the tag difference view shows $R_1\cap R_2$, $R_1/R_2$ and $R_2/R_1$
in addition to precision, recall and F-measure values. 
The user selects a predicate to compute the metrics from the following predicates:
(1) ``Intersection'': a tag from $R_1$ intersects in text with a tag in $R_2$,
(2) ``Exact'': a tag from $R_1$ exactly matches a tag in $R_2$,
(3) ``A includes B'': a tag from $R_1$ contains a tag from $R_2$, and
(4) ``B includes A'': a tag from $R_2$ contains a tag from $R_1$.

>>>>

%\framework saves the tags and the tag types in a user friendly format using
%the JavaScript Object Notation (JSON) format~\cite{nolan2014javascript}.
%The tag file keeps the paths of the separate text and tag type files. 
%It also contains a list of tags with their tag type identifiers,
%position in text, length, and word index.
%The tag type file contains MBF and MRE tag types. 
%%Each object has an array value containing the user-defined tag types. 
%Each tag type contains the name, description, expression, and the visualization legends.

@@

\begin{figure}[tb!]
  \centering
  \includegraphics[width=\textwidth]{figures/treegraph}
  \caption{MRE annotated Text, MRE \major{matching parse tree}, and entity-relation graph}
  \label{fig:treegraph}
\end{figure}

@@

The \major{matching parse tree} view shows the selected match in a graph view.
Figure~\ref{fig:treegraph} shows the \major{matching parse tree} of the direction task 
\RL{dby mwl `l_A mqrbT mn h_dA Almbn_A}(Dubai Mall is located near this building). 
%Leaf nodes show the match text, internal nodes show the operations, and edge labels shows the MBF tag type name.

@@

A snapshot of the GUI in 
Figure~\ref{fig:treegraph} shows in an interactive graph view
the entity-relation graph of the match of the user defined relation 
extracted from the \major{matching parse tree} of the MRE. 
%
In the computational action editor, an advanced user can 
enter C++ code and use the \framework API to program and process 
subexpression matches. 
%RE matches In order to add computational actions to an MRE sub-expression, 
%the user selects to edit the actions in the MRE tag type editor. 
%The view allows the user to specify the actions in a user friendly manner. 
%\framework provides the user with an API to access the sub-expression match features easily and define the pre-match actions. 
%The match features include the text, position, length, and numerical value if applicable, and morphological features for MBF matches. 
%\framework also enables the user to add C++ declarations and library includes.

@@

In the analysis view, the user provides 
two tag sets $R_1$ and $R_2$ and 
two tag type sets ${\cal T}_1$ and ${\cal T}_2$ as input. 
%The analysis view provides views of the differences between the tag types and 
%the tag sets.
%
The tag type difference view shows the text annotated in three panes: 
(i) the common tag types ${\cal T}_1 \cap {\cal T}_2$,
(ii) the tag types in ${\cal T}_1$ but not in ${\cal T}_2$, 
and (iii) the tag types in ${\cal T}_2$ and not in ${\cal T}_1$.
%
Similarly, the tag difference view shows $R_1\cap R_2$, $R_1/R_2$ and $R_2/R_1$
in addition to precision, recall and F-measure values. 
The user selects a predicate to compute the metrics from the following predicates:
(1) ``Intersection'': a tag from $R_1$ intersects in text with a tag in $R_2$,
(2) ``Exact'': a tag from $R_1$ exactly matches a tag in $R_2$,
(3) ``A includes B'': a tag from $R_1$ contains a tag from $R_2$, and
(4) ``B includes A'': a tag from $R_2$ contains a tag from $R_1$.

### 2018-04-20 tex/merf/TALMajor/introduction.tex
done till 3.19

@@

In this paper, 
we present \framework, a morphology-based entity and relational entity
extraction framework for Arabic text.
\framework provides a user-friendly interface where the user defines tag types 
and associates them with
%\framework formulae that are 
regular expressions over 
%\framework
Boolean formulae.
A Boolean formula is defined by a term, negation of a term, or disjunction of terms.
Terms are matches to Arabic morphological features including 
prefix, stem, suffix, part of speech (POS) tags, gloss tags, extended synonym tags, and semantic categories.
We often refer to regular expressions and Boolean formulae as expressions and formulae, respectively, for short.
%We discuss the importance of the morphological features supported in 
%\framework terms in Section~\ref{sec:morph}; in brief, morphological preprocessing
%is key to Arabic NLP.% due to the richness of Arabic morphology.

We illustrate the target of \framework using a simple example.
Consider the text in Figure~\ref{fig:intromotiv} which contains
directions to Dubai Mall~\footnote{text taken from the Dubai Mall website
~\url{http://www.thedubaimall.com/ar/}}.
\major{
The figure also presents a transliteration and an English translation of 
the Arabic text.
}
The framed words in the text are entities referring to names of people 
($n_1,n_2,n_3$), 
names of places ($p_1,\dots,p_7$), 
relative positions ($r_1,\dots,r_4$), 
and numerical terms ($u_1,u_2$). 
We would like to extract those entities, and then extract
the relational entities forming the graph in Figure~\ref{fig:intromotiv} 
where vertices express entities, 
and edges represent the relational entities.
\framework allows a regular user to do that interactively
in four phases that include morphological analysis, 
entity extraction based on morphological features, 
relational entity extraction, and entity cross-reference.

\framework regular expressions support operators such as concatenation, 
zero or one, zero or more, one or more, up to $M$ \major{repititions where $M$ is 
a non-zero positive integer}, and logical conjunction and disjunction.
An editor allows the user to associate an action with each sub-expression.
The user specifies the action with C++ code and uses an API to access information related to the matches such as text, position, length, morphological features, and numerical value.

\framework takes an Arabic text, a set of user-defined Boolean 
formulae, and a set of user-defined regular expressions.
\framework computes the morphological solutions of the words in the 
input text then computes matches to the Boolean formulae.
\framework then generates a {\em non-deterministic finite state 
automata} (NDFSA) for each expression and simulates it with the 
sequence of Boolean formulae matches to compute the regular 
expression matches. 
\framework generates executable code for the actions associated with
the regular expressions, 
compiles, links, and executes the generated code 
as shared object libraries.
Finally, \framework constructs the semantic relations and 
cross-reference between entities.
\framework also provides visualization tools to present the matches, 
and estimate their accuracy with respect to reference tags.

>>>>

\major{
Figure~\ref{fig:intromotiv} illustrates the target of \framework using 
the directions to Dubai Mall example~\footnote{text taken from the Dubai 
Mall website ~\url{http://www.thedubaimall.com/ar/}}.
The figure also presents a transliteration and an English translation of 
the Arabic text.
}
The framed words in the text are entities referring to names of people 
($n_1,n_2,n_3$), 
names of places ($p_1,\dots,p_7$), 
relative positions ($r_1,\dots,r_4$), 
and numerical terms ($u_1,u_2$). 
We would like to extract those entities, and then extract
the relational entities forming the graph in Figure~\ref{fig:intromotiv} 
where vertices express entities, 
and edges represent the relational entities.

@@

In this paper, 
we present \framework, a morphology-based entity and relational entity
extraction framework for Arabic text.
\framework provides a user-friendly interface where the user defines tag types 
and associates them with
%\framework formulae that are 
regular expressions over 
%\framework
Boolean formulae.
%
A Boolean formula is defined by a term, negation of a term, or disjunction of terms.
Terms are matches to Arabic morphological features including 
prefix, stem, suffix, part of speech (POS) tags, gloss tags, extended synonym 
tags, and semantic categories.
\major{
For example, entity $p_1$ in Figure~\ref{fig:intromotiv} has a place semantic 
category. 
}
\framework regular expressions support operators such as concatenation, 
zero or one, zero or more, one or more, up to $M$ \major{repititions where $M$ is 
a non-zero positive integer}, and logical conjunction and disjunction.
\major{
For example, the sequence between $p_1$ and $p_2$ matches a regular expression $re$ 
that requires two semantic place categories with a place-preposition POS 
tag ($r_1$) in between. 
}

An editor allows the user to associate an action with each sub-expression.
The user specifies the action with C++ code and uses an API to access information 
related to the matches such as text, position, length, morphological features, 
and numerical value.
%
\major{
Each regular expression is associated with a named identifier to form a 
{\em local grammar} like structure~\cite{traboulsi2009arabic}. 
%
A relation definition GUI allows the user to provide relational tuples 
where each tuple has a 
source, a destination and an edge label.
The user uses the regular expression identifiers to define the relational 
tuple elements.
For example, the relation between $e_1, e_2$ and $r$ shown in 
Figure~\ref{fig:intromotiv} is a match of a relational tuple 
over the components of $re$. 
%
We refer to regular expressions and Boolean formulae as expressions and formulae, respectively.
We also refer to expressions as rules
when used in a grammar context; e.g.  when used with an identifier. 
%We discuss the importance of the morphological features supported in 
%\framework terms in Section~\ref{sec:morph}; in brief, morphological preprocessing
%is key to Arabic NLP.% due to the richness of Arabic morphology.
}

\framework takes an Arabic text and the local grammar defined by the 
Boolean formulae and the regular expressions.
\framework computes the morphological solutions of the 
input text then computes matches to the Boolean formulae therein.
\framework then generates a {\em non-deterministic finite state 
automata} (NDFSA) for each expression and simulates it with the 
sequence of Boolean formulae matches to compute the regular 
expression matches. 
\framework generates executable code for the actions associated with
the regular expressions, 
compiles, links, and executes the generated code 
as shared object libraries.
Finally, \framework constructs the semantic relations and 
cross-reference between entities.
\framework also provides visualization tools to present the matches, 
and estimate their accuracy with respect to reference tags.

### 2018-04-20 tex/merf/TALMajor/methodology.tex
done till 3.19

The set of tag types ${\cal T}$ contains tuples of the form $\langle l,f,d\rangle$ 
where $l$ is a text label with a descriptive name, 
$f$ is an MRE, and $d$ is a visualization legend 
with font and color information.

For each word $t_i\in T, 0\le i < M, M=|T|$.
\framework computes a Boolean value 
($\{\mathit{true}, \mathit{false}\}$)
for all MBFs. 
Then, it computes the set of MBF tags
$R_i=\{(t_i,tt)| tt=\langle l,f,d\rangle \wedge
f~\mathit{is~an~MBF} \wedge f(t_i)\} \subseteq T \times {\cal T}$
which tags a word $t_i$ with $\mathit{tt}$ 
iff the MBF $f$ associated with
tag type $\mathit{tt}$ is true for $t_i$. 
The MBF evaluation results in a sequence of tag sets 
$\langle R_0, R_1, \ldots, R_{n-1}\rangle$.
If a word $t_i$ has no tag type match, 
its tag set $R_i$ is by default the singleton $O=\{\mathit{NONE}\}$.
%and $t_i$ is referred to as a {\em null} word.

%\subsection{MRE and action simulation}

For each MRE, 
\framework generates its equivalent non-deterministic finite automaton (NFA) in the typical manner~\cite{sipser2006introduction}. 
%Each MRE operation has its equivalent representation in an NFA. 
We support the upto operation ($f$\^{}$x$), which is not directly 
supported in ~\cite{sipser2006introduction}, by 
%it is equivalent to $f?|ff|\dots|\underbrace{f}_{x \text{ times}}$ which has an NFA mapping. 
%we can expand it into a standard regular expression form, for example
expanding it into a regular expression form; for example 
$f$\^{}$3$ is equivalent to $f?|ff|fff$. 
Consider the example of Figure~\ref{fig:intromotiv} and the
corresponding expression $(P|N)\!+~O?~R~O^\wedge 2~(P|N|U)+$. 
Figure~\ref{fig:nfaEx} shows part of the corresponding NFA where
$q_8, q_9, \dots, q_{13}$ represent NFA states,
and edges are transitions based on MBF tags such as 
$P,$ and $N$.
Edges labeled with the empty string $\epsilon$ are non-deterministic.
%The expression uses operations $|$, $?$, $+$, and $\wedge$ to relate places {\tt P}, names of persons {\tt N}, relative positions {\tt R}, numerical terms {\tt U}, and other {\tt O}.

\framework simulates the generated NFA over the sequence of tag sets matching the MBF formulae.
A simulation match $m$ of an expression $f$ is a tree where the root is the 
MRE expression, the internal nodes are the MRE and MBF operations, and the 
leaves are matches of the MAT terms of $f$.
The leave matches form a vector of tags
$\langle r_k,r_{k+1},\dots,r_j\rangle$ 
corresponding to the text sequence 
$\langle t_k,t_{k+1},\dots,t_j\rangle$ where $r_{\ell}\in R_{\ell},0\le k\le \ell \le j < n$.
%
%If the simulation has a match $\langle r_m,r_{m+1},\dots,r_n\rangle$ 
%where $0\le m\le n$, 
%the next simulation starts at $R_{n+1}$. 
%This disallows overlap of matches for the same MRE. 
%
%In case the NFA simulation has no match, 
%the next simulation starts at $R_{m+1}$. 
If we have more than one match %starting at $R_k$ where $0\le k\le n$, 
\framework returns the longest. 
%
Figure~\ref{fig:motiv}(b) shows two match trees of $\mathit{dir}$ extracted 
from the text of Figure~\ref{fig:intromotiv}.
\RL{dby} and \RL{mwl} are leaf nodes referring to name of place tags ($P$). 
The $+$, $|$, and $?$ MRE operations
are internal nodes.

%\framework maintains a function $\phi\subset Q\times\Phi$, 
%where $\mathit{Q}$ is the set of states in the NFA and $\Phi$ is the set of sub-expressions in the MRE.
%So $(q,f)\in \phi$ iff state $q\in Q$ was generated by \framework to correspond to sub-expression $f\in \Phi$. 
%\framework uses $\phi$ to compute a match tree with respect to the MRE regular expression. 
%It also uses $\phi$ and the match sequence to compute the sequence of computational actions of an MRE match.

\setcode{utf8}
\setarab
\transfalse

\begin{figure}[tb!]
\centering
\resizebox{0.7\columnwidth}{!}{
	\relsize{-1} \input{figures/reNFA.tex}
}
  \caption{Equivalent NFA of direction expression}
  \label{fig:nfaEx}
\end{figure}
\transtrue
\setcode{standard}

%\begin{figure}[tb!]
%{ \relsize{-1}
%\begin{framed}
%\begin{multicols}{2}
%\begin{itemize}
%\item Annotated Expression: \\
%\\
%{ \relsize{-2.5}
%%\begin{itemize}
%%  \item 
%  $\stackrel{e_1}{(P|N)+}~\stackrel{o_1}{\mathit{O}?}~\stackrel{r}{R}~\stackrel{o_2}{\mathit{O^{\wedge}2}}~\stackrel{e_2}{(P|N|U)+}$
%%\end{itemize}
%}
%\end{itemize}
%\begin{itemize}
%\item User defined semantic \\relations:
%\begin{itemize}
%\item $\langle e_1,e_2,r\rangle$
%\item $\langle r,e_1,o_1\rangle$
%\item $\langle r,e_2,o_2\rangle$
%\end{itemize}
%\end{itemize}
%\columnbreak
%\setcode{utf8}
%\transfalse
%\resizebox{0.9\columnwidth}{!}{
%	\relsize{+1.5} \input{figures/ergraphEx.tex}
%}
%\transtrue
%\setcode{standard}
%\end{multicols}
%\end{framed}
%}
%\caption{User-defined semantic relation example}
%\label{fig:srEx}
%\end{figure}

\framework computes the relational entities in a user defined 
relation $R=\ldbrack \langle e_1,e_2,r\rangle \rdbrack$
$\subseteq \ldbrack e_1 \rdbrack \times \ldbrack e_2 \rdbrack \times \ldbrack r \rdbrack$ 
to be the elements of $\ldbrack e_1 \rdbrack \times \ldbrack e_2 \rdbrack \times \ldbrack r \rdbrack$
with the smallest nonzero positive distance between the source and the destination
where the distance is the number of words between the matches.

%\framework supports the following configurations for the subexpressions of $e_i$ and $e_j$:
%\begin{itemize}
%\item $\left\vert{e_i}\right\vert=1\wedge\left\vert{e_j}\right\vert=1$: Relate $e_i$ match to $e_j$ match with edge labelled by $r$.
%\item $i=j\wedge\left\vert{e_i}\right\vert>1$: Relate sequences of $e_i$ matches with edges labelled by $r$.
%\item $i\neq j\wedge\left\vert{e_i}\right\vert>1\wedge\left\vert{e_j}\right\vert=1$: Relate each $e_i$ match to the single $e_j$ match with an edge labelled by $r$. The reverse applies for $\left\vert{e_i}\right\vert=1\wedge\left\vert{e_j}\right\vert>1$.
%\item $i\neq j\wedge\left\vert{e_i}\right\vert>1\wedge\left\vert{e_j}\right\vert>1$: Relate each $e_i$ match to each $e_j$ match with an edge labelled by $r$.
%\end{itemize}
%
In Figure~\ref{fig:motiv}(b), \framework names the subexpressions 
$(P|N)+$, $(P|N|U)+$, $O?$, $O\wedge 2$, and $R$, 
as $e_1, e_2, o_1, o_2,$ and $r$, respectively. 
The user defines the semantic relations
$\langle e_1, e_2, r\rangle$, 
$\langle r, e_1, o_1\rangle$, and 
$\langle r,e_2,o_2\rangle$.

The matches of $e_1,~e_2,~o_1,~o_2,$ and $r$ from the second match tree 
of Figure~\ref{fig:motiv}(b) are \RL{dby mwl}(Dubai Mall), 
\RL{Almbn_A}(the building), \RL{`l_A}(prep), \RL{mn h_dA}(from this), 
and \RL{mqrbT}(near), respectively. 
\framework~constructs the semantic relation matches and builds the lower part of the 
entity-relation graph shown in Figure~\ref{fig:intromotiv}.

For the first match \RL{brj _hlyfT bAlqrb mn AltqA.t` Al-'-wl},
the matches of $e_1,~e_2,~o_2,$ and $r$ are 
\RL{brj _hlyfT}(Khalifa tower), \RL{AltqA.t` Al-'-wl}(intersection 1), 
\RL{mn}(from), and \RL{bAlqrb}(next to), respectively. 
\framework~doesn't construct the relation $\langle r, e_1, o_1\rangle$ 
since $o_1$ has no match. 
Therefore, we get the upper part of the entity-relation graph shown 
in Figure~\ref{fig:intromotiv}.

After computing the relational entities, 
\framework~computes a cross-reference relation between the extracted entities
using a second order synonymy feature ($Syn^2$).
The \cci{isA} edge in the 
graph of Figure~\ref{fig:intromotiv} shows the cross-reference relation 
between \RL{brj _hlyfT}(Khalifa tower) from the first match 
with \RL{Almbn_A}(the building) from the second match.

>>>>

Figure~\ref{f:overview} illustrates the four processes involved in 
\framework methodology. 
The first process takes the Arabic text input and provides the user 
with a morphology-based Boolean (MB) formulae GUI. 
The user interactively composes MB-formulae using the GUI and the output 
of the simulator and the $Syn^k$ detector. 
The simulator and the detector apply the formulae over the morphological solutions 
of the Arabic text and produce the MB-formulae tags.

The second process takes the MB-formulae tags and the Arabic text and provides
the user with a morphology-based grammar rule GUI. 
The user interactively composes MB-grammar rules using the GUI and the output 
of the MB-grammar rule simulator. 
The grammar rule simulator applies the rules over the MB-formulae tags and 
produces the MB-grammar rule tags.

The third process takes the MB-grammar rule tags and provides the user with 
a relation and action GUI. 
The user interactively provides (1) the relation definitions and 
(2) the actions in terms of identifiers from the MB-grammar 
rules.
The relation extraction produces the target entities and relational
entities. 
The action execution enriches the entities and the relational entities
with powerful semantics.
For example, users can utilize actions to compute statistical features, 
store intermediate results, 
or apply intelligent entity inference techniques
as we show later in the numerical extraction example of 
Subsection~\ref{subsec:numnorm}.
%
Finally, in the fourth process the user compares the results with golden
reference chunks and visualizes the difference. 
This allows the user to refine the formulae, rules, relations and actions.

After the relation construction phase, we are interested to relate the 
discovered entities that express the same concept.
\framework provides the extended synonym feature of second order as 
a default cross-reference relation ($Syn^2$).
In Figure~\ref{fig:intromotiv}, triggering this feature creates the 
edge labels with \cci{isSyn} 
between the nodes \cci{Khalifa Tower} and \cci{The building}.

\input{MERFOverview}

The user may inspect and refine the defined formulae, rules and relations and 
the corresponding formulae tags, rule tags, entities and relational entities
either using the GUI or directly through readable output files. 
The files are in the javascript object notation (JSON)~\cite{nolan2014javascript} 
format that is intuitive to read and modify. 
\framework separates the user defined formulae, rules, actions and relations in
a \framework tag type file and the 
matching tags in a tags files. 
The separation serves the user to apply the tag types to multiple case studies 
and to obtain a separate resulting tags file for each case study.
